A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-16 latex/main.tex
Updates from Overleaf

To compare the performance of sc2ts and UShER+RIPPLES 
in terms of recombinant detection, we ran RIPPLES on the Viridian UShER phylogeny
subset to the 2,475,418 samples shared with the sc2ts ARG (STAR Methods).
Using the RIPPLES parameter $p=4$ (equivalent to $k=4$ used 
to build the sc2ts ARG) resulted in 1174 reported recombination events.
We ran 3SEQ on each of the candidate trios (STAR Methods) and 601 
were classified as recombinant, giving an overall validation rate of 51\%.
Aligning the events reported by both methods using the sets of descendant
samples (STAR Methods), 468 of the events are shared by both, 
with the intersection comprising 40\% of the RIPPLES events
and 55\% of the sc2ts events (Figure~\ref{fig:ripples_sc2ts_events}). 
The 706 recombination events output by RIPPLES that do not 
correspond to a recombination in the sc2ts ARG can largely
be explained by a combination of two factors.
Firstly, 408 of these RIPPLES identified recombination events
have one or more parents that are \emph{younger} than the
recombinant sequence based on metadata. Thus, the parental
sequences were simply not available to match against for 
sc2ts (which enforces strict time ordering) when the 
recombinant sequence was inserted. Secondly, of the 179
remaining events with matching clades and in which the 
sc2ts node has more than 4 mutations, of the total
1,167 mutations, 452 of these occurred at sites that were 
remapped during post-hoc parsimony. The remaining ``missing''
events are likely attributable to the finer details of the 
differences in tree building approaches between UShER
and sc2ts.

>>>>

To compare the performance of sc2ts and UShER+RIPPLES 
in terms of recombinant detection, we ran RIPPLES on the Viridian UShER phylogeny
subset to the 2,475,418 samples shared with the sc2ts ARG (STAR Methods).
Using the RIPPLES parameter $p=4$ (equivalent to $k=4$ used 
to build the sc2ts ARG) resulted in 1174 reported recombination events.
We ran 3SEQ on each of the candidate trios (STAR Methods) and 601 
were classified as recombinant, giving an overall validation rate of 51\%.
Aligning the events reported by both methods using the sets of descendant
samples (STAR Methods), 468 of the events are shared by both, 
with the intersection comprising 40\% of the RIPPLES events
and 55\% of the sc2ts events (Figure~\ref{fig:ripples_sc2ts_events}). 
The 706 recombination events output by RIPPLES that do not 
correspond to a recombination event in the sc2ts ARG can largely
be explained by a combination of two factors.
Firstly, 408 of these RIPPLES identified recombination events
have one or more parents that are \emph{younger} than the
recombinant sequence based on metadata. Thus, the parental
sequences were simply not available to match against for 
sc2ts (which enforces strict time ordering) when the 
recombinant sequence was inserted. Secondly, of the 179
remaining events with matching clades and in which the 
sc2ts node has more than 4 mutations, of the total
1,167 mutations, 452 of these occurred at sites that were 
remapped during post-hoc parsimony. The remaining ``missing''
events are likely attributable to the finer details of the 
differences in tree building approaches between UShER
and sc2ts.
